---
title: "p8105_hw3_yl4924"
author: "Yongzheng Li"
date: "2022-10-15"
output: github_document
---

# Load the package adjust graph display
```{r, message=FALSE,warning=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 2

### Part 1: load, clean and tidy the data
```{r}
accel_data <- read_csv(
              "./data/accel_data.csv") %>% 
              janitor::clean_names() %>% 
              pivot_longer(
                starts_with("activity"),
                names_to = "minute",
                names_prefix = "activity_",
                values_to = "activity_count") %>% 
              mutate(
                minute = as.integer(minute),
                type_of_day = case_when(
                  day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ 1,
                  day %in% c("Saturday", "Sunday")                                   ~ 2),
                type_of_day = recode_factor(type_of_day, `1` = "weekday", `2` = "weekend"),
                day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday", "Sunday")))
```

The accel_data contain `r nrow(accel_data)` number of observations of a 63 year old male with bmi 25 in 35 days. The dataset contain `r ncol(accel_data)` variables. The variables include week, day_id, day, minute, activity_count, and day_of_week. Among these variables, activity_count is the most important variable. The mean activity_count of the male in 35 days is `r mean(accel_data$activity_count)` per minutes.

### Part 2: total activity over the day
```{r}
accel_data %>% 
  group_by(week, day) %>% 
  summarize(total_day_count = sum(activity_count)) %>% 
  pivot_wider(
    names_from = "day",
    values_from = "total_day_count") %>% 
  knitr::kable(digits = 1)

```

From the output table, there is a increasing trend of activity count on weekdays in week 1, week 2, and week 5. Week 4 has a decreasing trend of activity count on weekdays. It is worth to note that On Saturday in week 4 and 5, the total activity count is 1440 that means there is only 1 activity count per minute in these two days. The value is probably a missing value and we should give more attention to these values.

### Part 3: single-panel plot that show 24 hour activity time courses 
```{r}
accel_data %>% 
ggplot(aes(x = minute, y = activity_count, color = day)) +
  geom_line(alpha = 0.7) +
  labs(
    title = "24-hour activity time courses for each day",
    x = "Minute",
    y = "Activity Count"
  )
```
From the plot, minutes from 0 to 250 has relative low activity count(below 1250 counts). The activity count increased from 250 mins to 1125 mins(around 1875 counts except Sunday). The activity count increased to the highest from 1125 mins to 1300 mins(above 3000 counts). The activity count decreased below 1250 counts after 1300 mins(around 1250 counts).

# Problem 3

### load the view "ny_noaa" data 
```{r}
data("ny_noaa")
head(ny_noaa)
skimr::skim(ny_noaa)
```

from the the summary result, the data contain 2595176 observations and 9 variables. The key variables are prcp, snow, snwd, tmax, and tmin. There are 145838, 381221, and 591786 missing values in variable prcp, snow and snwd that account 5.6%, 14.7%, 22.8% percent of values in each variable, respectively. There are also missing values in variables tmax and tmin 

### bullet point 1: data cleaning
Clean date by seperating date into year, month, day, change prcp unit to mm, convert tmax, tmin into numerical variable and change unit to degrees C.
```{r}
ny_noaa <- ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), "-") %>% 
  mutate(
    tmax = as.numeric(tmax)*0.1,
    tmin = as.numeric(tmin)*0.1,
    prcp = prcp*0.1,
    month = month.name[as.integer(month)]
  )

find_mode <- function(x) {
  u <- unique(x)
  tab <- tabulate(match(x, u))
  u[tab == max(tab)]
}
```

The most commonly observed values for snowfall(variable snow) is `r find_mode(ny_noaa$snow)` mm. The reason that 0 mm snowfall is the most observed value is because snow only falls when temperate is very low(around 0 degree c). In every year, only some of the countries can have snowfalls during the winter. Therefore, it is usual to see the most observed value of snowfall = 0. 

### bullet point 2: 
```{r}
ny_noaa %>% 
  filter(month %in% c("January", "July")) %>% 
  group_by(id, month, year) %>% 
  summarize(average_max_temp = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = average_max_temp, group = id)) +
  geom_line(alpha = 0.7) +
  facet_grid(. ~ month)
  
```

From the plot, it is clear to see that average maximum temperature in July is greater than January. The average maximum temperate fluctuate across years. The plot shows the pattern that the average maximum temperate increase for 2-3 years follow by a decrease for 2-3 years. There are some outliers for both month. One of the outliers has an average maximum temperate of -13 degrees of C in January in 1982.

### bullet point 3:

since the plot of tmax vs tmin for the full dataset has large number of data, I choose to use hex plot to plot tmax vs tmin. for plot 2, boxplot is helpful to show the distribution for different variables. 
```{r}
tmax_tmin_plot <- 
  ny_noaa %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex(alpha = 0.7)

dist_snow_plot <-
  ny_noaa %>% 
  filter(0 < snow & snow < 100) %>% 
  ggplot(aes(x = snow, y = year)) +
  geom_boxplot()

tmax_tmin_plot + dist_snow_plot

```



